Seismic Hazard Estimation of Northern Iran Using Smoothed Seismicity

### 2017-01-26 attenuation.tex
completed changes based on reviewer 3 comments and our response

The selected backbone GMPE for our analysis, as proposed by \citet{Kalkan2004}, is of the form
%
\begin{align}
	\ln \left( Y \right) =
		& \hspace{1ex} b_1 + b_2(M_w - 6) + b_3 \left( M_w - 6 \right)^{2} \nonumber \\
		& + b_5 \ln \left( r \right) + b_V \ln \left( V_S / V_A \right)
	\, ,
\end{align}
%
where $Y$ is a ground motion parameter (here representing PGA), $V_A$ is a reference velocity in m/s, and $V_S$ is the shear wave velocity of the site of interest. In this equation, the distance $r$ is given by
%
\begin{equation}
	r= \sqrt{ r^2_{\mathit{cl}} + h^2 }
	\, ,
\end{equation}
%
where $r_{\mathit{cl}}$ is the horizontal distance to the site of interest and $h$ is a reference fictitious depth, both given in km. According to \citet{Kalkan2004}, in the case of $Y$ representing PGA, the coefficients $b_1$, $b_2$, $b_3$, $b_5$ and $b_V$ are
%
\begin{equation}
\begin{array}{lcrlcr}
	b_1 &=&  0.393   \,,&\hspace{2em}   b_2 &=& 0.576\,,   \\
	b_3 &=& -0.107   \,,&\hspace{2em}   b_5 &=& -0.899\,,  \\
	b_V &=& -0.200   \,;
	\nonumber
\end{array}
\end{equation}
%
and $V_A$, $V_S$ and $h$ are
%
\begin{equation}
\begin{array}{lcrl}
	V_A &=& 1,112 & \mathrm{m/s}	\\
	V_S &=&   700 & \mathrm{m/s}	\\
	h   &=&  6.91 & \mathrm{km}\,,
	\nonumber
\end{array}
\end{equation}
%
respectively. The standard deviation of the residuals $(\sigma_{\ln y})$ expressing the random variability of the ground motions is 0.612. The value of $V_S$ is assumed to represent average surface rock sites.

@@

We should note here that even though \citet{Kalkan2004} derived the above attenuation relationship for distances $r$ up to 250 km, \citet{Zafarani2014} only used data up to 200 km in their evaluation of the various GMPEs available. Therefore, to be consistent with the latter, we set the models in the hazard analysis to compute ground motion intensities (i.e., PGAs) at distances no greater than 200 km.

>>>>

The selected backbone GMPE for our analysis, as proposed by \citet{Kalkan2004}, is of the form
%
\begin{align}
	\ln \left( Y \right) =
		& \hspace{1ex} b_1 + b_2(M_w - 6) + b_3 \left( M_w - 6 \right)^{2} \nonumber \\
		& + b_5 \ln \left( r \right) + b_V \ln \left( V_S / V_A \right)
	\, ,
\end{align}
%
where $Y$ is a ground motion parameter (here representing PGA), $V_A$ is a reference velocity in m/s, and $V_S$ is the shear wave velocity of the site of interest. In this equation, the distance $r$ is given by
%
\begin{equation}
	r= \sqrt{ r^2_{\mathit{cl}} + h^2 }
	\, ,
\end{equation}
%
where $r_{\mathit{cl}}$ is the horizontal distance to the site of interest and $h$ is a reference fictitious depth, both given in km. \myrevision{We should note here that given the fact that the smoothed seismicity method assumes all earthquakes as individual point sources, $r_{\mathit{cl}}$ is the same as the epicentral distance.} According to \citet{Kalkan2004}, in the case of $Y$ representing PGA, the coefficients $b_1$, $b_2$, $b_3$, $b_5$ and $b_V$ are
%
\begin{equation}
\begin{array}{lcrlcr}
	b_1 &=&  0.393   \,,&\hspace{2em}   b_2 &=& 0.576\,,   \\
	b_3 &=& -0.107   \,,&\hspace{2em}   b_5 &=& -0.899\,,  \\
	b_V &=& -0.200   \,;
	\nonumber
\end{array}
\end{equation}
%
and $V_A$, $V_S$ and $h$ are
%
\begin{equation}
\begin{array}{lcrl}
	V_A &=& 1,112 & \mathrm{m/s}	\\
	V_S &=&   700 & \mathrm{m/s}	\\
	h   &=&  6.91 & \mathrm{km}\,,
	\nonumber
\end{array}
\end{equation}
%
respectively. The standard deviation of the residuals $(\sigma_{\ln y})$ expressing the random variability of the ground motions is 0.612. The value of $V_S$ is assumed to represent average surface rock sites.

@@

We should note here that even though \citet{Kalkan2004} derived the above attenuation relationship for distances $r$ up to 250 km, \citet{Zafarani2014} only used data up to 200 km in their evaluation of the various GMPEs available. Therefore, to be consistent with the latter, we set the models in the hazard analysis to compute ground motion intensities (i.e., PGAs) at distances no greater than 200 km. \myrevision{We should also note that the values of $V_A$, $V_S$ and $h$ were kept constant throughout the calculations for this study. This was done because these are the values for which the attenuation relationship showed good agreement with PGA data from northern Iran at rock sites. While it would be tempting to adjust these values to the specific site conditions of each grid point, or to match the hypocentral depth of each earthquake in the catalog, doing this would effectively modify the goodness-of-fit of the GMPE with the data, and thus make the predictions inaccurate.}

### 2017-01-26 compilation.tex
completed changes based on reviewer 3 comments and our response

For the particular case of our region of interest, the catalog prepared by \citet{Zare2014} included events $M_w \geq 4$ between 550 B.C.~and 2006. We took information from the catalog up to December 1999, and augmented it by adding newer and more complete data from IIEES for events $M \geq 3$ between January 2000 and December 2015. While we recognize that earthquakes in the range $3 \leq M_w \leq 4$ are unlikely to cause structural damage, we include them as a means to feed the model with information about epicentral locations with potential for future larger magnitude earthquakes \citep{Kafka_2000_BSSA}, and to help inform the Guten\-berg-Richter relationship. By contrast, in the recent seismic hazard analysis done for Iran by \citet{Khodaverdian_2016_BSSA}, the authors used a similar catalog for events $M_w \geq 4$, but only included data through 2011, thus excluding the 2012 $M_w$ 6.4 Tabriz earthquake, which severely damaged the northeast part of the country.

>>>>

For the particular case of our region of interest, the catalog prepared by \citet{Zare2014} included events $M_w \geq 4$ between 550 B.C.~and 2006. We took information from the catalog up to December 1999, and augmented it by adding newer and more complete data from IIEES for events $M \geq 3$ between January 2000 and December 2015 \myrevision{(no formal catalog exists with earthquakes $M_w \leq 4$ prior to 2000)}. While we recognize that earthquakes in the range $3 \leq M_w \leq 4$ are unlikely to cause structural damage, we include them as a means to feed the model with information about epicentral locations with potential for future larger magnitude earthquakes \citep{Kafka_2000_BSSA}, and to help inform the Guten\-berg-Richter relationship. By contrast, in the recent seismic hazard analysis done for Iran by \citet{Khodaverdian_2016_BSSA}, the authors used a similar catalog for events $M_w \geq 4$, but only included data through 2011, thus excluding the 2012 $M_w$ 6.4 Tabriz earthquake, which severely damaged the northeast part of the country.

### 2017-01-26 conclusions.tex
completed changes based on reviewer 3 comments and our response

We present results from a seismic hazard analysis for northern Iran based on a smoothed seismicity approach, including regional maps of expected peak ground accelerations for different seismic zonation models. In particular, we calculated the hazard for the region of interest using two basic models, a five-seismic zone regions model and a uniform seismic zone model, including uncertainty in the combination of both models, the seismic parameters, and the selection of a ground motion prediction equation via a logic tree approach. In general, our calculations yield results similar to those of other studies for the region of interest and the main cities in it. We, however, provide a newer and alternative perspective to the problem of seismic hazard analysis for norther Iran than previous studies because our use of a more complete seismic catalog, which includes recent damaging events and low-magnitude earthquakes. In our analysis we found that the combination of the regional and uniform models not only allowed us to account for the uncertainty in seismic zonation but it led to a final hazard map that adjusted better to the visual interpretation of the historical and instrumental seismicity observed in the region of interest, the location of known seismically active faults, and the low-magnitude seismic activity in the region. While no one model can be deemed more appropriate than others, we believe our results constitute a positive contribution to the understanding of seismic hazard in the region, especially for rural areas that are not the subject of specialized hazard studies such as those conducted for major urban centers. Future work to improve the results presented here should consider other ground motion measures of intensity such as response spectral accelerations for different periods and site effects.

>>>>

We present results from a seismic hazard analysis for northern Iran based on a smoothed seismicity approach, including regional maps of expected peak ground accelerations for different seismic zonation models. In particular, we calculated the hazard for the region of interest using two basic models, a five-seismic zone regions model and a uniform seismic zone model, including uncertainty in the combination of both models, the seismic parameters, and the selection of a ground motion prediction equation via a logic tree approach. In general, our calculations yield results similar to those of other studies for the region of interest and the main cities in it. We, however, provide a newer and alternative perspective to the problem of seismic hazard analysis for norther Iran than previous studies because our use of a more complete seismic catalog, which includes recent damaging events and low-magnitude earthquakes. In our analysis we found that the combination of the regional and uniform models not only allowed us to account for the uncertainty in seismic zonation but it also led to a final hazard map that adjusted better to the visual interpretation of the historical and instrumental seismicity observed in the region of interest, the location of known seismically active faults, and the low-magnitude seismic activity in the region. While no one model can be deemed more appropriate than others, we believe our results constitute a positive contribution to the understanding of seismic hazard in the region, especially for rural areas that are not the subject of specialized hazard studies such as those conducted for major urban centers \myrevision{or specific sites of interest near notable faults with a history of damaging earthquakes}. Future work to \myrevision{complement and} improve the results presented here should consider other ground motion measures of intensity such as response spectral accelerations for different periods and site effects\myrevision{, as well as alternative ground motion prediction equations suitable for northern Iran. All things considered, we recognize that additional efforts are necessary to obtain a more robust estimate of hazard, especially for specific areas of interest. In this regard, future efforts should be directed to the combination of studies like this for the contribution of the background seismicity with more targeted studies considering extended source models for areas near major fault lines. It should also be noted that this is not a final assessment of the seismic hazard of the region, and decisions regarding seismic design or risk mitigation plans should consider this as well as other studies.}

### 2017-01-26 introduction.tex
completed changes based on reviewer 3 comments and our response

Over the last three decades there have been multiple efforts to estimate the seismic parameters and seismic hazard of Iran, and improve the assessment of seismic risk in the country \citep[e.g.,][]{Tavakoli1999, Moinfar_2000_Chap, Ghodrati2003, Moinfar_2012_WCEE, Khodaverdian_2016_BSSA}. Altogether, these previous studies have helped improve the Iranian Code of Practice for Seismic Resistant Design of Buildings \citep{BHRC2014}. Such improvements are reflected, for instance, in lower overall expected damage ratios, as reported by \citet{Ghodrati2013} for the city of Tehran. There is, however, room for improvement. The most recent of these studies, \citet{Khodaverdian_2016_BSSA}, presents an estimation of spatially-variable seismic parameters for Iran using an approach similar to the one we use here, but does not necessarily focus on hazard. Other previous efforts have outdated catalogs, use fault-centered hazard analyses, or focus on particular locations or cities.

>>>>

Over the last three decades there have been multiple efforts to estimate the seismic parameters and seismic hazard of Iran, and improve the assessment of seismic risk in the country \citep[e.g.,][]{Tavakoli1999, Moinfar_2000_Chap, Ghodrati2003, Moinfar_2012_WCEE, Khodaverdian_2016_BSSA}. Altogether, these previous studies have helped improve the Iranian Code of Practice for Seismic Resistant Design of Buildings \citep{BHRC2014}. Such improvements are reflected, for instance, in lower overall expected damage ratios, as reported by \citet{Ghodrati2013} for the city of Tehran. There is, however, room for improvement. The most recent of these studies, \citet{Khodaverdian_2016_BSSA}, presents an estimation of spatially-variable seismic parameters for Iran using an approach similar to the one we use here, but does not necessarily focus on hazard\myrevision{, and does not include the latest available data for the region}. Other previous efforts \myrevision{also} have outdated catalogs, use fault-centered hazard analyses, or focus on particular locations or cities. \myrevision{By contrast, our analysis presents a regional perspective which is likely to be more useful for rural areas that are usually not the subject of specialized studies on hazardous faults}.

### 2017-01-26 new-response.tex
completed changes based on reviewer 3 comments and our response

This being said, we did not want to let this opportunity pass by without also addressing the points raised by the third reviewer. We were a little intrigued by the strict criticism of this third reviewer, but at the same time understood the motivation behind his/her comments. While we appreciated very much your candid reaction this this review---given our response to the initial evaluation---, we thought that in addition to following your advise about how to handle it (through minor modifications of form and language), it was still appropriate for us to share with you our views about some of these criticisms. You will also find enclosed our responses, which we hope you will appreciate.

@@

\comment{1}{%
As shown in Figure 2 of this manuscript, the major faults and its characteristicparameter uncertainties should be incorporated into the seismic hazard study of the this region. As the authors stated in Section 4, this type of work has been performed in Tavakoli and Ghafory-Ashtiany (1999) and many others. In order to improve the previous studies, therefore, the seismic hazard maps presented in this manuscript should be modified based on the known major fault sources and areal seismic sources, which are considered as background seismic sources. The smoothed seismicity approach could be used for the background seismic source zones. Otherwise, only using the smoothed seismicity approach could lead to misleading of seismic hazard values and hazard maps in this region, especially when the time span of earthquake catalog is not enough to give us the true seismicity pattern of the region. Lack of consideration for major faults, will have significant and far-reaching consequences for Iran and major metropolitan areas such as Tehran. Accepting the publication in a reputable journal such as Journal of Seismology without consideration for major faults will be a mistake. The reviewer is puzzled why the fourth author of this article who is a reputable researcher has agreed on the publication of this article. The authors write: ``We find our results to be helpful in understanding seismic hazard for northern Iran, but recognize that additional efforts are necessary to obtain more robust estimates at specific areas of interest and different site conditions.''
}

@@

>>>>

This being said, we did not want to let this opportunity pass by without also addressing the points raised by the third reviewer. We were a little intrigued by the strict criticism of this third reviewer, but at the same time understood the motivation behind his/her comments. While we appreciated very much your candid reaction to this third review---given our response to the two initial ones---, we thought that in addition to following your advise about how to handle it (through minor modifications of form and language), it was still appropriate for us to share with you our views about some of these criticisms. You will also find enclosed our responses, which we hope you will appreciate. The latest changes to the manuscript are highlighted in the annotated version.

@@

\vspace{3ex}
\comment{1}{%
As shown in Figure 2 of this manuscript, the major faults and its characteristicparameter uncertainties should be incorporated into the seismic hazard study of the this region. As the authors stated in Section 4, this type of work has been performed in Tavakoli and Ghafory-Ashtiany (1999) and many others. In order to improve the previous studies, therefore, the seismic hazard maps presented in this manuscript should be modified based on the known major fault sources and areal seismic sources, which are considered as background seismic sources. The smoothed seismicity approach could be used for the background seismic source zones. Otherwise, only using the smoothed seismicity approach could lead to misleading of seismic hazard values and hazard maps in this region, especially when the time span of earthquake catalog is not enough to give us the true seismicity pattern of the region. Lack of consideration for major faults, will have significant and far-reaching consequences for Iran and major metropolitan areas such as Tehran. Accepting the publication in a reputable journal such as Journal of Seismology without consideration for major faults will be a mistake. The reviewer is puzzled why the fourth author of this article who is a reputable researcher has agreed on the publication of this article. The authors write: ``We find our results to be helpful in understanding seismic hazard for northern Iran, but recognize that additional efforts are necessary to obtain more robust estimates at specific areas of interest and different site conditions.''
}

@@

Changes in the manuscript motivated by the points raised by the reviewer are highlighted in the annotated version of the manuscript, which reflect our view towards as expressed in the responses above.
